---
title: "Coherence-Linkages"
output: html_document
date: "2023-04-18"
---

# Summary
Data for this code are available at: https://doi.org/10.7910/DVN/VK3CP/
This code explains the analysis carried out for assessing coherence of information in NDCs and NAPs. 
Specifically, it helps to generate Figure 3 in the Manuscript illustrating the coherence between adaptation cycle elements. 
In the figure, nodes represent the six adaptation elements. Links indicate the frequency of co-occurrence between two elements. n represents the total number of countries that identify a type of adaptation element (hazards, system at risk, goal, objective, action, or indicator); or that establish linkages between two pairs of elements. The wider the link, the more countries explicitly establish a link between the corresponding pair of elements. Each color represents an adaptation element. The grey circle highlights four stages of the adaptation cycle: assessment, planning, implementation, and monitoring and evaluation. 

which illustrates a he chord diagram representing how 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(circlize)
library(ggplot2)
```
# Preparation
To run this code, download the file "AdaptationElementsLinkages" available at: https://doi.org/10.7910/DVN/VK3CP.
First, you have to first create an adjancency matrix. This is explained in the R Mark Down file: "0.AdaptationElementsLinkages_Processed"
The code below starts by uploading the matrix from the local file. 
```{r}
links <- read.csv(file = "C:/Users/anowak/OneDrive - CGIAR/Desktop/Repositories/adaptation-tracking-africa/Data-Processed/LinkagesMatrix.csv", header = TRUE)
#links_full <- read.csv(file = "Data-Raw/LinkagesRaw.csv", header = TRUE)
rownames(links) <- make.unique(links$X)
links$X <- NULL
# rename Systems at risk
colnames(links)[2] <- "System at risk"
rownames(links)[2] <- "System at risk"
```

# Figure 3
Create a symetric matrix, ensuring  the values on the upper triangle of the matrix match the corresponding values on the lower triangle. 
Therefore, use lower.tri() function to identify the lower triangle of the matrix, and then assign the corresponding values from the transposed matrix using t(symmetric_matrix). After making these changes, the matrix will be symmetric matrix. The diagonal elements are left unchanged, assuming they represent the connection of a group to itself (which should typically be zero). 
You can adjust the figure to include legends and annotations. 
```{r}
# create matrix
mylinks <- as.matrix(links)

lower_triangular <- lower.tri(mylinks)
mylinks[upper.tri(mylinks)] <- 0
print(mylinks)

# Set colors for the plot
set.seed(999)
colors <- c("Hazard" = "#c7eae5", 
            "System at risk" =  "#5ab4ac",
            "Goal" = "#01665e",
            "Objective" =  "#8c510a", 
            "Action" = "#d8b365",
            "Indicator" = "#f6e8c3")

# Create an image to store the graphic data. Adjust the file path to 
png("C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Results/Fig3.png", width = 4, height = 4, units = "in", res = 330, bg = "transparent") 

# Plot
circos.par(gap.after = 10)
circos.par(start.degree = 20) #to start with Hazards 
par(mar = c(1, 1, 1, 1))  # Adjust the margins
par(cex = 0.8)
chordDiagram(mylinks, column.col = colors, grid.col = colors, transparency = 0.3, annotationTrack = "grid", annotationTrackHeight = mm_h(1))

# Clear circular layout parameters
circos.clear() 
dev.off()
```
