---
title: "ED_ElementsVsLinkages"
output: html_document
date: "2023-10-26"
---

# Summary
Data for this code are available at: https://doi.org/10.7910/DVN/VK3CP, files "AdaptationElementsLinkages" and "AdaptationElements"
This code generates a scatter plot showing the relationship between comprehensiveness and coherence (Extended Data Figure 3). 
Specifically, it illustrates the relationship between number of elements and number of linkages between elements in a country (NDC and NAP combined). Each dot represents an African country (N=53). The blue regression line indicates the linear relationship between the two variables. Country names are expressed using their ISO 3166-1 alpha-3 codes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(ggplot2)
library(dplyr)
library(ggrepel)  # For non-overlapping text labels
```

# Preparation
DonloElements vs Linkages
To run this code, download the files "AdaptationElements" and "AdaptationElementsLinkages" available at: https://doi.org/10.7910/DVN/VK3CP.
Adjust the file path below with the correct location where you saved the files

First, you will have to create a table that shows: country, n linkages in a country (based on "AdaptationElementsLinkages" file) and n elements and n element types in a country (based on "AdaptationElements" file). Then call the table as a new dataframe.
"Elem_links" represents the new dataframe generated, which has been stored on a local file.  
```{r}
elem_links <- read.csv(file = "C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Data/Data-Processed/ElementsVsLinkages.csv", fileEncoding = "Windows-1252")
```

# Extended data Figure 3: Relationship betwee
```{r}
elem_links$n_linkages <- as.numeric(elem_links$n_linkages) 
elem_links$n_elements <- as.numeric(elem_links$n_elements) 
elem_links$n_element_types <- as.numeric(elem_links$n_element_types)
```
Plot
```{r}
elem_links_scatter <- ggplot(elem_links) +
  geom_point(aes(x=n_linkages, y=n_elements)) + 
  geom_text_repel(aes(x=n_linkages, y=n_elements, label=iso3), size=3) +
  geom_smooth(aes(x=n_linkages, y=n_elements), method="lm", col="#8bcac4") + 
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),      # Removing major grid lines
    panel.grid.minor = element_blank(),      # Removing minor grid lines
    axis.line = element_line(size = 0.5),    # Adding thin lines for axes
    axis.ticks = element_line(size = 0.5)) +   # Adding thin ticks
  labs(title="", x="n linkages", y="n elements")
elem_links_scatter
```
## Analyse correlation
```{r}
# calculate Pearsons correlation coefficient
correlation <- cor(elem_links$n_linkages, elem_links$n_elements)
print(correlation)

# test significance of the correlation
cor_test <- cor.test(elem_links$n_linkages, elem_links$n_elements)
print(cor_test)
```

## Save data to local files
Modify the file path to specify the location where you would like to store the new dataframe
```{r}
ggsave("C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Results/ED_Fig3.png", elem_links_scatter, width=9, height=5)
```

