---
title: "NAPvsNDC-Comprehensiveness-Coherence-Robustness"
output: html_document
date: "2023-11-09"
---

# Summary
Data for this code are available at: https://doi.org/10.7910/DVN/VK3CP
This code explains the analysis carried out for comparing comprehensiveness, coherence and robustness of NDCs and NAPs. 
Specifically, it helps to generate individual panels for Figure 1a-c in the Extended Data.

Comprehensiveness is defined by the number of elements (n) featured in each country document (min. 3 and max. 6 elements);
Coherence is defined by the number of linkages (n) between the six adaptation elements featured in each country document (min. 1 and max. 15 linkages);
Robustness is defined by the number of SMART criteria (n) met by adaptation indicators in each country document (min.1 and max. 5 criteria). The assessment focuses on the best performing indicators and therefore provides an aggregate snapshot of overall robustness of the information in a document. If a document's adaptation indicators met different criteria to different extents, we used the maximum number of criteria fulfilled as the specific document's benchmark. SMART criteria refers to: Specific, Measurable, Achievable, Relevant, and Timebound. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyr)
library(dplyr)
```

# Preparation
To run this code, download the files "AdaptationElements" and "AdaptationElementsLinkages" available at: https://doi.org/10.7910/DVN/VK3CP.
Adjust the file path below with the correct location where you saved the files. 
```{r}
adaptdata <- read.csv(file = "C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Data/Data-Processed/AdaptationElementsProcessed.csv", fileEncoding = "Windows-1252")
links_full <- read.csv(file = "C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Data/Data-Raw/AdaptationElementsLinkages.csv", header = TRUE)
indicators <- adaptdata[adaptdata$Element == "Indicator", ]
```
# Extended Data Figure 1
## Comprehensiveness
Prepare
```{r}
df <- adaptdata[, c("Country", "Document", "Element")]

df_summary <- df %>%
  group_by(Country, Document) %>%
  summarise(Comprehensiveness = n_distinct(Element)) %>%
  ungroup()

# Filter data for NDC and NAP
ndc_data_com <- df_summary[df_summary$Document == "NDC", ]
nap_data_com <- df_summary[df_summary$Document == "NAP", ]

breaks_com <- c(0, 1, 2, 3, 4, 5, 6)
```
Plot
```{r}
png("C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Results/ED_Figure1a.png", width = 3, height = 3, units = "in", res = 330)

par(cex = 0.7)

hist_com <- {
  hist(ndc_data_com$Comprehensiveness, breaks = breaks_com, main = "", xlab = "Elements (n)", ylab = "Countries (n)", col = "#01665e", ylim = c(0, 25), xaxt = "n")
  mean_ndc_com <- mean(ndc_data_com$Comprehensiveness)
  sd_ndc_com <- sd(ndc_data_com$Comprehensiveness)
  curve(dnorm(x, mean = mean_ndc_com, sd = sd_ndc_com) * length(ndc_data_com$Comprehensiveness) * diff(breaks_com)[1], col = "#001a18", lwd = 2, add = TRUE)
}

hist(nap_data_com$Comprehensiveness, breaks = breaks_com, main = "", xlab = "Elements (n)", ylab = "Countries (n)", col = rgb(214, 179, 101, alpha = 128, maxColorValue = 255), add = TRUE, ylim = c(0, 30), xaxt = "n")
mean_nap_com <- mean(nap_data_com$Comprehensiveness)
sd_nap_com <- sd(nap_data_com$Comprehensiveness)
curve(dnorm(x, mean = mean_nap_com, sd = sd_nap_com) * length(nap_data_com$Comprehensiveness) * diff(breaks_com)[1], col = "#8c510a", lwd = 2, add = TRUE)

ndc_counts_com <- hist(ndc_data_com$Comprehensiveness, breaks = breaks_com, plot = FALSE)
nap_counts_com <- hist(nap_data_com$Comprehensiveness, breaks = breaks_com, plot = FALSE)

#legend("topright", legend = c("NDC", "NAP"), fill = c("#01665e", rgb(214, 179, 101, alpha = 128, maxColorValue = 255), x = "topright", y = NULL))

dev.off()
```
## Coherence
Prepare
```{r}
# Convert empty spaces to 0, then convert columns 5:19 to numeric
links_full[, 5:19] <- sapply(links_full[, 5:19], function(x) as.numeric(ifelse(x == "", 0, x)))

# Handle potential NA values after conversion, replacing with 0
links_full[, 5:19][is.na(links_full[, 5:19])] <- 0

# Add the total_links column
links_full$total_links <- rowSums(links_full[, 5:19])

# subset
links_full <- links_full[, c("Country", "Document", "total_links")]

links_summary <- links_full

# Filter the data for NDC and NAPs
ndc_data_coh <- links_summary[links_summary$Document == "NDC", ]
nap_data_coh <- links_summary[links_summary$Document == "NAP", ]
 
breaks_coh <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)
```
Plot
```{r}
png("C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Results/ED_Figure1b.png", width = 3, height = 3, units = "in", res = 330)

par(cex = 0.7)

hist_coh <- {
  hist(ndc_data_coh$total_links, breaks = breaks_coh, main = "", xlab = "Linkages (n)", ylab = "Countries (n)", col = "#01665e", ylim = c(0, 10), xaxt = "n")
  mean_ndc_coh <- mean(ndc_data_coh$total_links)
  sd_ndc_coh <- sd(ndc_data_coh$total_links)
  curve(dnorm(x, mean = mean_ndc_coh, sd = sd_ndc_coh) * length(ndc_data_coh$total_links) * diff(breaks_coh)[1], col = "#001a18", lwd = 2, add = TRUE)
}

hist(nap_data_coh$total_links, breaks = breaks_coh, main = "", xlab = "Linkages (n)", ylab = "Countries (n)", col = rgb(214, 179, 101, alpha = 128, maxColorValue = 255), add = TRUE, ylim = c(0, 10), xaxt = "n")
mean_nap_coh <- mean(nap_data_coh$total_links)
sd_nap_coh <- sd(nap_data_coh$total_links)
curve(dnorm(x, mean = mean_nap_coh, sd = sd_nap_coh) * length(nap_data_coh$total_links) * diff(breaks_coh)[1], col = "#8c510a", lwd = 2, add = TRUE)

ndc_counts_coh <- hist(ndc_data_coh$total_links, breaks = breaks_coh, plot = FALSE)
nap_counts_coh <- hist(nap_data_coh$total_links, breaks = breaks_coh, plot = FALSE)

#legend("right", legend = c("NDC", "NAP"), fill = c("#01665e", rgb(214, 179, 101, alpha = 128, maxColorValue = 255)))

dev.off()
```
# Robustness (indicators)
```{r}
# <- indicators[, c(2, 3, 20:24)]

indicators_new <- indicators %>%
  mutate_at(vars(20:24), ~ifelse(. == "yes", 1,
                               ifelse(. == "no", 0, as.numeric(.))))

indicators_new2 <- indicators_new %>%
  group_by(Country, Document) %>%
  mutate(nr_criteria = Specific + Measurable + Achievable + Relevant + Timebound) 
  #summarise(nr_criteria = sum())

indicators_new3 <- indicators_new2[, c(2, 3, 25)] 

indicators_new3 <- indicators_new2 %>%
  group_by(Country, Document) %>%
  summarise(nr_criteria = max(nr_criteria, na.rm = TRUE))

# Filter the data for NDC and NAPs
ndc_data_r <- indicators_new3[indicators_new3$Document == "NDC", ]
nap_data_r <- indicators_new3[indicators_new3$Document == "NAP", ]
 
breaks_r <- c(0, 1, 2, 3, 4, 5)
```
Plot
```{r}
png("C:/Users/anowak/OneDrive - CGIAR/PhD/Ch1/Results/ED_Figure1c.png", width = 3, height = 3, units = "in", res = 330)

par(family = "Arial", cex = 0.7)

hist_r <- {
  hist(ndc_data_r$nr_criteria, breaks = breaks_r, main = "", xlab = "SMART criteria (n)", ylab = "Countries (n)", col = "#01665e", ylim = c(0, 10), xaxt = "n")
  mean_ndc_r <- mean(ndc_data_r$nr_criteria)
  sd_ndc_r <- sd(ndc_data_r$nr_criteria)
  curve(dnorm(x, mean = mean_ndc_r, sd = sd_ndc_r) * length(ndc_data_r$nr_criteria) * diff(breaks_r)[1], col = "#001a18", lwd = 2, add = TRUE)
}

hist(nap_data_r$nr_criteria, breaks = breaks_r, main = "", xlab = "SMART criteria (n)", ylab = "Countries (n)", col = rgb(214, 179, 101, alpha = 128, maxColorValue = 255), add = TRUE, ylim = c(0, 10), xaxt = "n")
mean_nap_r <- mean(nap_data_r$nr_criteria)
sd_nap_r <- sd(nap_data_r$nr_criteria)
curve(dnorm(x, mean = mean_nap_r, sd = sd_nap_r) * length(nap_data_r$nr_criteria) * diff(breaks_r)[1], col = "#8c510a", lwd = 2, add = TRUE)

ndc_counts_r <- hist(ndc_data_r$nr_criteria, breaks = breaks_r, plot = FALSE)
nap_counts_r <- hist(nap_data_r$nr_criteria, breaks = breaks_r, plot = FALSE)

#legend("top", legend = c("NDC", "NAP"), fill = c("#01665e", rgb(214, 179, 101, alpha = 128, maxColorValue = 255)))

dev.off()
```
